While looking for related work, Erin and I came across \cite{harsh2020place}, which is cited in \cite{ingrosso2022data}.
This paper uses a restricted Boltzmann machine (RBM) to model hippocampal place cells.
They sample data from an Ising model to train the RBM.
In particular, they show analytically that the RBM learns localized RFs.
This is the only work I've come across that explicitly models place cells, and I think this could be a more relevant and exciting application of our work.
It seems that the question of why place cells emerge is still open, and I think our work could be a step in the right direction.

\cite{harsh2020place} just focuses on the Ising model and doesn't try to make general claims about which distributions will lead to place-cell emergence.
They connect their findings to previous work on ``retarded learning,'' a general phenomenon in which a minimum of data is needed for a phenomenon (in this case, place cells) to emerge.
This doesn't seem like a direction we can readily tackle, since we've been operating in the full-batch setting.

\cite{harsh2020place} recalls Gardner's theory of optimal learning for a single-layer perceptron \cite{gardner1988space}, which predicts that data drawn from a translation-invariant distribution will lead to the emergence of localized RFs.
At a first glance this seems to make a (false) prediction, but upon closer reading it actually seems to just be making a prediction about the Ising model.
Nevertheless, \cite{harsh2020place} make it sound like translation-invariant data is a sufficient condition for place-cell emergence, which is not true.
It seems like our work, which draws a sharper distinction, might be a useful contribution here.

\subsection*{Summary}
\emph{I have not found much other work on modeling place cells, but I will keep looking.}
This seems like an underexplored area, and I think it could be a good application of our work.
Our toy model is much closer to existing work in this space than for work on V1.
Additionally, localization and place-cell emergence are basically the same thing, whereas V1 RFs have the additional complexity of orientation and phase selectivity that I have not been able to address yet.
