% GO TO: /Users/leonlufkin/Library/texmf/tex/latex
%        to add custom packages to the path
\documentclass{article}
\usepackage{report}

\title{Literature Review \& Summary}
\author{Leon Lufkin}
\date{\today}

\makeatletter
\let\Title\@title
\let\Author\@author
\let\Date\@date

\begin{document}

\section{Internship Summary}
\label{sec:internship_summary}
The goal of my internship was to come up with a normative model for the emergence of the structure shown in Alessandro's paper \cite{ingrosso2022data}.
There were three phenomena: splitting into two populations of neurons, tiling, and the emergence of localized receptive fields (RFs).
The first phenomenon was relatively easy to explain, and the second is quite challenging mathematically but makes a lot of intuitive sense.
The third phenomenon, the emergence of localized RFs, is the most intriguing and the least understood.
It is especially intriguing because of its apparent connection to the development of localized, oriented edge detectors in primary visual cortex, and perhaps also place cells in hippocampus.

Here, I will explore to what extent we could explain that phenomenon using some of the analysis I did during my internship.
First, I will summarize my work on understanding localization.
Then, I will summarize existing models of visual cortex development.
Following that, I will summarize existing work on the emergence of place cells.
Finally, I will discuss whether it makes sense to try to connect the two, and if so, how we might do that.

\section{Localization}
\label{sec:localization}
\input{localization.tex}

\section{V1 RFs}
\label{sec:v1_rfs}
\input{v1_rfs.tex}

\section{Place Cells}
\label{sec:place_cells}
\input{place_cells.tex}

%% Works Cited
% \newpage
\bibliographystyle{alpha}
\bibliography{refs}


\end{document}
